Performance-portable Parallel Programming in Computer Science Education with Kokkos

### 2020-01-24 src/3.MachineModel.tex
Adding other example and back-end example.
Updated img.

The Kokkos abstract machine representation targets a design of a future shared-memory computing architectures. The design is shown in Figure~\ref{fig:execspace}. It is characterized by multiple latency-optimized execution units and off-die bandwidth-optimized accelerators. Both compute device types can have disjoint memory address spaces with unique performance properties. In such an architecture, execution units might be hierarchically organized with multiple levels of achievable parallelism with different memory access characteristics and coherence properties of caches. In order to ensure performance portability on such a wide range of configurations, an abstraction of compute resources and available memories are required. For this purpose we introduce the abstract concept called \emph{space}.

@@

An instance of a memory space provides a concrete method for the application programmer to request data storage allocations. Following the architecture as discussed earlier, the multi-core processor may contain multiple memory spaces that abstract on-package memory, slower DRAM and non-volatile memories. Accelerators can provide an additional memory space through its local on-package memory. An example of such an architectural layout is shown in Figure~\ref{fig:memspace}. The programmer is free to decide where each data structure is allocated by requesting the corresponding memory space. Programmatically this is achieved by instantiating the according memory space. Kokkos provides the appropriate abstraction of memory allocation and data management. We believe that an abstract representation of execution- and memory devices is a key property towards performance-portable parallel programming.

@@

\begin{figure}
\begin{subfigure}[b]{0.41\textwidth}
\centering
\includegraphics[width=\textwidth]{img/ExecSpaces.png}
\caption{Execution space is an instantiation of a space that represents a processing device to which the programmer can target parallel code.}
\label{fig:execspace}
\end{subfigure}
\hfill \break
\begin{subfigure}[b]{0.41\textwidth}
\centering
\includegraphics[width=\textwidth]{img/MemSpaces.png}
\caption{A memory space is an instantiation of a space that represents a memory location on which the programmer can allocate data.}
\label{fig:memspace}
\end{subfigure}
\caption{Spaces represent conceptual building blocks of the abstract machine model in Kokkos.}
\label{fig:spaces}
\end{figure}

>>>>

The Kokkos abstract machine representation targets a design of a future shared-memory computing architectures. The design is shown in Figure~\ref{fig:KokkosSpaces}. It is characterized by multiple latency-optimized execution units and off-die bandwidth-optimized accelerators. Both compute device types can have disjoint memory address spaces with unique performance properties. In such an architecture, execution units might be hierarchically organized with multiple levels of achievable parallelism with different memory access characteristics and coherence properties of caches. In order to ensure performance portability on such a wide range of configurations, an abstraction of compute resources and available memories are required. For this purpose we introduce the abstract concept called \emph{space}.

@@

An instance of a memory space provides a concrete method for the application programmer to request data storage allocations. Following the architecture as discussed earlier, the multi-core processor may contain multiple memory spaces that abstract on-package memory, slower DRAM and non-volatile memories. Accelerators can provide an additional memory space through its local on-package memory. This is highligted in Figure~\ref{fig:KokkosSpaces} in a green color. shows memory spaces in . The programmer is free to decide where each data structure is allocated by requesting the corresponding memory space. Programmatically this is achieved by instantiating the according memory space. Kokkos provides the appropriate abstraction of memory allocation and data management. We believe that an abstract representation of execution- and memory devices is a key property towards performance-portable parallel programming.

@@

\begin{figure}

\includegraphics[width=0.5\textwidth]{img/KokkosSpaces.png}
\caption{Spaces represent conceptual building blocks of the abstract machine model in Kokkos.}
\label{fig:execspace}
\label{fig:spaces}
\end{figure}
